Finiteness and Cardinality in Homotopy Type Theory

### 2020-01-07 paper.tex
down to 16 pages

  We formalise our work in Cubical Agda
  \cite{vezzosiCubicalAgdaDependently2019}, and we implement a library for proof
  search (including combinators for level-polymorphic fully generic currying),
  and demonstrate how it can be used to both prove properties and synthesise
  full functions given desired properties.
\end{abstract}
\section{Introduction}
In this work, we will explore finite types in Cubical Type Theory
\cite{cohenCubicalTypeTheory2016}, and expose their relationship to infinite
types, species, and demonstrate their practical uses for proofs in dependently
typed programming languages.
\subsection{Strong Finiteness Predicates}
We will first explore the ``strong'' notions of finiteness (i.e. those at least
as strong as Kuratowski finiteness \cite{kuratowskiNotionEnsembleFini1920}),
with a special focus on cardinal finiteness (section~\ref{cardinal}), and
manifest enumerability (section~\ref{manifest-enumerability}), which is new, to
our knowledge.

\input{figures/finite-classification}
Figure~\ref{finite-classification} organises the predicates according to
their ``strength''; i.e. how much information they provide about a conforming
type.
For instance, a proof that some type \(A\) is manifestly Bishop finite (the
strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
us that \(A\) is discrete (has decidable equality), and gives us a linear order
on the type.
A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
features: indeed, we will see examples of Kuratowski finite types which are not
even sets, never mind discrete ones.

@@

For each predicate, we will also prove its closure over sums and products in
both dependent and non-dependent forms, if such a closure exists.
This will culminate in our main result for this section: the formal proof that
decidable Kuratowski finite sets form a topos.
More specifically, we show that cardinal finite sets form a topos, that
decidable Kuratowski finite sets are equivalent in strength to cardinal finite
sets, and carry the proof over.
This result relies on proofs on each of the other finiteness predicates.

\subsection{Infinite Types}
In section~\ref{infinite-cardinalities}, we will extend our study of finite
types to infinite but countable types.
We will see that the classification of finiteness predicates mirrors that of
their countable counterparts, and we will prove closure under the Kleene star.

@@

Before giving the definition of the predicate, we will first define lists (and
membership thereof): we have chosen a \emph{container}-based definition for this
work.
\begin{rm-definition}[Container] \label{container-def}
  A container \cite{abbottContainersConstructingStrictly2005} is a pair
  \(S \triangleright P\) where \(S\) is a type, the elements of which are called
  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
  the elements of \(P(s)\) are called the \emph{positions} of a container.
  We ``interpret'' a container into a functor defined like so:
  \begin{equation}
    \llbracket S \triangleright P \rrbracket(A) = \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
  \end{equation}
  Membership of a container can be defined like so:
  \begin{equation}
    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
  \end{equation}
  Where \(\text{fiber}\) is from \cite[definition 4.2.4]{hottbook}.
\end{rm-definition}
\begin{rm-definition}[Lists]
  \begin{equation}
    \mathbf{List} = \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket
  \end{equation}
\end{rm-definition}
\begin{rm-definition}[\mathbf{Fin}]
  \(\mathbf{Fin}\)(n) is the type of natural numbers smaller than \(n\).
  It is defined inductively, as follows:
  \begin{equation}
    \begin{aligned}
      \mathbf{Fin}(0) &= \bot \\
      \mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)
    \end{aligned}
  \end{equation}
\end{rm-definition}

@@

Finally, we can define formally split enumerability.
\begin{rm-definition}[Split Enumerable Set]
    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in xs
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
\subsection{Split Surjections}
\begin{rm-definition}[Surjections]
  \begin{multicols}{2}
      A \twoheadrightarrow B = \Sigma (f : A \rightarrow B) , \text{surjective}(f)
    \end{equation} \break
      A \twoheadrightarrow! \; B = \Sigma (f : A \rightarrow B) , \text{split surjective}(f)
  \end{multicols}
  Where the definitions of surjective are taken from \cite[definition
  4.6.1]{hottbook}.
An equivalent way to define ``finiteness'' is via a split surjection from a
finite prefix of the natural numbers.
\begin{rm-theorem} \label{split-enum-surj}
  Split enumerability is equivalent to a split surjection from a finite prefix
  of the natural numbers.
    \mathcal{E}!(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  The proof is surprisingly short: after sufficient inlining, it emerges that
  our goal is simply a reassociation.
  \ExecuteMetaData[agda/Cardinality/Container/Finite.tex]{split-surj}
  To be clear: in Agda, the proof could simply be \(\AgdaFunction{reassoc}\); we
  have written out the extra lines for clarity alone.
\subsection{Decidable Equality}
\begin{rm-lemma} \label{split-enum-discrete}
  Any split enumerable type has decidable equality (is discrete).
  We use a corollary that if there is a split-surjection from \(A\) to \(B\),
  and \(A\) is discrete, then \(B\) is also discrete.
By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, we know that
this implies split enumerable types must all be sets.
In this section we will prove closure under various operations for split
enumerable sets.
We are working towards a topos proof, which requires us to prove closure under
a variety of operations: for now, we only have enough machinery to demonstrate
the semiring operations, and dependent sums.
In order to show closure under exponentials (function arrows), we will need an
equivalence with \(\mathbf{Fin}\), which will be provided in
section~\ref{manifest-bishop}.
  \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
  These non-recursive types have similar, simple proofs.
  \begin{multicols}{3}
    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool}
    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{unit}
    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{empty}
  \end{multicols}
\begin{rm-theorem} \label{split-enum-sigma}
  Split-enumerability is closed under \(\Sigma\).
  \begin{equation} \label{split-enum-sigma-clos-sig}
    \frac{
      \mathcal{E}!(A) \; \; \; \Pi (x : A) , \mathcal{E}!(U(x))
    }{
      \mathcal{E}!(\Sigma (x : A) , U(x))
    }
  To obtain the support list, we concatenate the support lists of all the proofs
  of split-finiteness for \(U\) over the support list of \(E_A\).
  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-enum}
  ``do-notation'' is available to us as we're working in the list monad.
Closure under disjoint union and Cartesian product both follow from \(\Sigma\),
as these types can be defined in terms of \(\Sigma\).
\begin{align}
  A \times B &= \Sigma {(x : A)} , B \\
  A + B &= \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
\end{align}
\section{Manifest Bishop Finiteness} \label{manifest-bishop}
Where split enumerability was the enumeration form of a split surjection from
\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
\emph{equivalence} with \(\mathbf{Fin}\).

@@

A nice consequence of prohibiting duplicates is that now the length of the
support list is the same as the cardinality of the set.
\subsection{Equivalence}
\begin{rm-lemma} \label{bishop-equiv}
  A proof of manifest Bishop finiteness is equivalent to an equivalence with a
  finite prefix of the natural numbers.
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{rm-lemma}
\begin{proof}
  There are many equivalent definitions of equivalence in HoTT.
  Here we take the version preferred in the Cubical Agda library: contractible
  maps \cite[definition 4.4.1]{hottbook}.
  Because of the parallels between contractible maps and split surjections,
  the proof proceeds much the same as \ref{split-enum-surj}.
  In other words, the definition of Bishop finiteness is itself a reassociation
  of a contractible map.
\end{proof}
\subsection{Relationship to Split Enumerability}
We now show that manifest Bishop finiteness has equal strength to split
enumerability.
\begin{rm-theorem} \label{split-enum-to-manifest-bishop}
  Any split enumerable set is manifest Bishop finite.
\end{rm-theorem}
\begin{proof}
  Let \(E\) be a proof of split enumerability for some set \(A\).
  From proposition~\ref{split-enum-discrete} we can derive decidable equality on
  \(A\), and using this we can define a function \(\text{uniques}\) which filters
  out duplicates from lists of \(A\)s.
  \begin{equation}
    \text{uniques} : \textbf{List}(A) \rightarrow \textbf{List}(A)
  \end{equation}
  This gives us our support list.

  It suffices now to prove the following:
  \begin{equation}
    \Pi {(x : A)} , \Pi {(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
  \end{equation}
  And from that we can generate our cover proof.
\end{proof}
Note that while manifest Bishop finiteness and split enumerability are equivalent
in ``strength'', the two types are not equivalent.
For instance, for a type \(A\) with \(n\) inhabitants, there are infinitely many
inhabitants of \(\mathcal{E}!(A)\), while there are only \(n!\) inhabitants of
\(\mathcal{B}(A)\).
\subsection{Closure}
Proving equal strength of split enumerability and manifest Bishop finiteness
allows us to carry all of the previous proofs of closure over to manifest Bishop
finite sets (and vice-versa).
Missing from our previous proofs was a proof of closure of functions.
We remedy that here.
\begin{rm-theorem}
  Manifest bishop finiteness is closed over dependent functions
  (\(\prod\)-types).
  \begin{equation}
    \frac{
      \mathcal{B}(A) \; \; \; \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
    }{
      \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
    }
  \end{equation}
\end{rm-theorem}
\begin{proof}
  This proof is essentially the composition of two transport operations, made
  available to us via univalence.

@@

  Seen another way, a permutation is an isomorphism of cover proofs.

@@

We also prove some of the identities you might expect with regards to
permutations.

@@

Based on previous proofs, we can derive that if we add decidability to a
Kuratowski finite type we retrieve cardinal finiteness.
  Any Kuratowski finite set with decidable equality is cardinally finite.
  We first convert to a propositionally truncated proof of manifest
  enumerability.
  From here, we can convert to manifest Bishop finiteness under the truncation
  with our decidable equality proof.

@@

  Kuratowski finite sets are closed under \(\Sigma\).

@@

  This follows from the \(\Sigma\) closure proof on manifestly enumerable types.

@@

While the previous section purported to be about finite sets, we can now see
that it was really only studying surjections and isomorphisms of different
flavours between types and \(\mathbf{Fin}\).
The natural next question which arises, then, is if we can extend that work to
surjections and isomorphisms with \(\mathbb{N}\).
In more standard language, what we're referring to here is of course
countable infinity and related concepts.
However, as with finite types, the ``countably infinite'' types have more
varieties in constructive mathematics than their classical counterparts.

@@

have a support \emph{stream}, which is its infinite, coinductive counterpart.

@@

  We will work with two isomorphic definitions of streams.
  The first is the following:

@@

    \mathbf{Stream}(A) = \mathbb{N} \rightarrow A

@@

  Of course, such a type can not be defined in the same way as a list: it would
  be impossible to construct a value, as inductive types do not admit
  infinitely-sized inhabitants.

@@

We can now define the split countable types.

@@

  As mentioned, we will have to use a more sophisticated pairing function than
  the Cartesian product we used before.
  We instead will mirror the pattern in figure~\ref{cantor}.
  To greatly simplify the algorithm, we will produce an intermediate stream of
  lists which consists of the diagonals in the diagram.
  We then concatenate these streams into the final support stream.
\end{proof}
\begin{rm-lemma}
  Split countability is closed under non-dependent product and sum.
\end{rm-lemma}
\begin{proof}
  Follows from theorem~\ref{split-countability-sigma}.
\end{proof}
\subsubsection{Kleene Star}
While we lose some closures with the inclusion of infinite types, we gain some
others.
In particular, we have the Kleene star.
This means, in effect, that we have closure under lists.
\begin{rm-theorem}
  Split countability is closed under Kleene star.
  \begin{equation}
    \mathcal{E}!(A) \rightarrow \mathcal{E}!(\mathbf{List}(A))
  \end{equation}
\end{rm-theorem}
\begin{proof}
  As with the proof of closure under \(\Sigma\), our main task here is to figure
  out a way to arrange the indices such that 
\end{proof}
\subsection{Manifest Countability}
As we can quotient out the position information with finite types, so can we with
countable types.
\todo[inline]{fill in rest here}
\section{Practical Uses} \label{practical}
The theory of finite types in constructive mathematics, and in HoTT in
particular, is rich and interesting, as we hope we have demonstrated thus far.
As well as being theoretically interesting, however, the proofs and combinators
we have defined here are \emph{practically} useful.
\subsection{Omniscience}
In this section we are interested in restricted forms of the limited principle
of omniscience \cite{myhillErrettBishopFoundations1972}.
\begin{rm-definition}[Limited Principle of Omniscience]
  For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
  omniscience is as follows:
  \begin{equation}
    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Sigma {(x : A)} , P(x) \right)
  \end{equation}
  In other words, for any decidable predicate the existential quantification of
  that predicate is also decidable.
\end{rm-definition}
The limited principle of omniscience is non-constructive, but individual types
can themselves satisfy omniscience.
In particular, \emph{finite} types are omniscient.

@@

Omniscience is stronger than exhaustibility, as we can derive the latter from
the former:
\begin{rm-lemma} \label{omniscient-is-exhaustible}
  Any omniscient type is exhaustible.
\end{rm-lemma}
\begin{proof}
  For decidable propositions, we know the following:
  \begin{equation} \label{lem-conversion}
    \Pi (x : A) , P(x) \leftrightarrow \neg \Sigma (x : A) , \neg P(x) 
  \end{equation}
  To derive exhaustibility from omniscience, then, we run the predicate in its
  negated form, and then subsequently negate the result.
  The resulting decision over \(\neg \Sigma(x : A) , \neg P(x)\) can be
  converted into \(\Pi (x : A) , P(x)\).
\end{proof}
We cannot derive, however, that any exhaustible type is omniscient, as we do not
have the inverse of equation~\ref{lem-conversion}:
\begin{equation} \label{choice-conversion}
  \Sigma (x : A) , P(x) \leftrightarrow \neg \Pi (x : A) , \neg P(x)
\end{equation}
Such an equation would allow us to pick a representative element from any type,
which is therefore non-constructive.
In a sense, equation~\ref{lem-conversion} requires a form of LEM on the
proposition (i.e. requires it to be decidable), whereas
equation~\ref{choice-conversion} requires a form of choice.
Those finiteness predicates which are ordered do in fact give us this form of
choice, so the conversion is valid.
As such, all of the ordered finiteness predicates imply omniscience.
Again, we will prove it only for the weakest.
\begin{rm-theorem}
  Manifest enumerable types are omniscient.
\end{rm-theorem}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}
Finally, we do have a form of omniscience for prop-valued predicates, as they do
not care about the chosen representative.
\begin{rm-theorem}
  Kuratowski finite types are omniscient about prop-valued predicates.
\end{rm-theorem}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}
\subsection{Synthesising Pattern-Matching Proofs}
In particular, they can automate large proofs by analysing every possible case.
In \cite{firsovDependentlyTypedProgramming2015}, the \(\AgdaDatatype{Pauli}\)
group is used as an example.
\ExecuteMetaData[agda/Data/Pauli.tex]{def}
For this type, there are several practical tasks we hope to achieve with the
help of our finiteness combinators:
\begin{itemize}
  \item Define decidable equality on the type
  \item Define the group operation on the type
  \item Prove properties about the group operation
\end{itemize}
Unfortunately, the simple pattern-matching way to do many of these tasks is
prohibitively verbose.
As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases.
A proof of decidable equality is one such function: it can be seen fully
worked-through in the appendix.

The alternative is to derive the things we need from \(\Enum!\) somehow.
First, then, we need an instance for \(\AgdaDatatype{Pauli}\):
\ExecuteMetaData[agda/Data/Pauli.tex]{inst}
From here we can already derive decidable equality, a function which requires 16
cases if implemented manually.

For proof search, the procedure is a well-known one in Agda
\cite{devrieseBrightSideType2011}: we ask for the result of a decision procedure
as an \emph{instance argument}, which will demand computation during
typechecking.
\ExecuteMetaData[agda/Cardinality/Finite/Search.tex]{auto}
And finally we can prove the property we wanted to on \(\AgdaDatatype{Pauli}\)
like so:
\ExecuteMetaData[agda/Data/Pauli.tex]{noninst-prf}

>>>>

  We formalise our work in Cubical Agda
  \cite{vezzosiCubicalAgdaDependently2019}, and we implement a library for proof
  search (including combinators for level-polymorphic fully generic currying),
  and demonstrate how it can be used to both prove properties and synthesise
  full functions given desired properties.
\end{abstract}
\section{Introduction}
\subsection{Strong Finiteness Predicates}
We will first explore the ``strong'' notions of finiteness (i.e. those at least
as strong as Kuratowski finiteness \cite{kuratowskiNotionEnsembleFini1920}),
with a special focus on cardinal finiteness (section~\ref{cardinal}), and
manifest enumerability (section~\ref{manifest-enumerability}), which is new, to
our knowledge.

Figure~\ref{finite-classification} organises the predicates according to
their strength; i.e. how much information they provide about a conforming
type.
For instance, a proof that some type \(A\) is manifestly Bishop finite (the
strongest of the notions, explored in section~\ref{manifest-bishop}) also tells
us that \(A\) is discrete (has decidable equality), and gives us a linear order
on the type.
A type that is Kuratowski finite (section~\ref{kuratowski}) has no such extra
features: indeed, we will see examples of Kuratowski finite types which are not
even sets, never mind discrete ones.

\input{figures/finite-classification}

@@

For each predicate, we will also prove its closure over sums and products in
both dependent and non-dependent forms, if such a closure exists.
This will culminate in our main result for this section: the formal proof that
decidable Kuratowski finite sets form a topos.
\subsection{Infinite Types}
In section~\ref{infinite-cardinalities}, we will extend our study of finite
types to infinite but countable types.
We will see that the classification of finiteness predicates mirrors that of
their countable counterparts, and we will prove closure under the Kleene star.

@@

Before giving the definition of the predicate, we will first define lists (and
membership thereof): we have chosen a \emph{container}-based definition for this
work.
\begin{rm-definition}[Container] \label{container-def}
  A container \cite{abbottContainersConstructingStrictly2005} is a pair
  \(S \triangleright P\) where \(S\) is a type, the elements of which are called
  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
  the elements of \(P(s)\) are called the \emph{positions} of a container.
  We ``interpret'' a container into a functor defined like so:
  \begin{equation}
    \llbracket S \triangleright P \rrbracket(A) = \Sigma {(s : S)} , \left( P(s) \rightarrow A \right)
  \end{equation}
  Membership of a container can be defined like so:
  \begin{equation}
    x \in \mathit{xs} = \text{fiber}(\text{snd}(\mathit{xs}), x)
  \end{equation}
  Where \(\text{fiber}\) is from \cite[definition 4.2.4]{hottbook}.
\end{rm-definition}
\begin{rm-definition}[Lists]
  \begin{equation}%
    \mathbf{List} = \llbracket \mathbb{N} \triangleright \mathbf{Fin} \rrbracket
  \end{equation}
\end{rm-definition}
\begin{rm-definition}[\mathbf{Fin}]
  \(\mathbf{Fin}(n)\) is the type of natural numbers smaller than \(n\).
  We define it the standard way, where \(\mathbf{Fin}(0) = \bot\) and
  \(\mathbf{Fin}(n + 1) &= \top + \mathbf{Fin}(n)\).
\end{rm-definition}

@@

Finally, we can define formally split enumerability.
\begin{rm-definition}[Split Enumerable Set]
  \begin{equation}
    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in xs
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
\end{rm-definition}
\subsection{Split Surjections}
\begin{rm-definition}[Surjections]
  \begin{multicols}{2}
    \begin{equation}
      A \twoheadrightarrow B = \Sigma (f : A \rightarrow B) , \text{surjective}(f)
    \end{equation} \break
    \begin{equation}
      A \twoheadrightarrow! \; B = \Sigma (f : A \rightarrow B) , \text{split surjective}(f)
    \end{equation}
  \end{multicols}
  Where the definitions of surjective are taken from \cite[definition
  4.6.1]{hottbook}.
\end{rm-definition}
\begin{rm-theorem} \label{split-enum-surj}
  Split enumerability is equivalent to a split surjection from a finite prefix
  of the natural numbers.
  \begin{equation}
    \mathcal{E}!(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{rm-theorem}
\begin{proof}
  The proof is surprisingly short: after sufficient inlining, it emerges that
  our goal is simply a reassociation.
  \ExecuteMetaData[agda/Cardinality/Container/Finite.tex]{split-surj}
  To be clear: in Agda, the proof could simply be \(\AgdaFunction{reassoc}\); we
  have written out the extra lines for clarity alone.
\end{proof}

\begin{rm-lemma} \label{split-enum-discrete}
  Any split enumerable type has decidable equality (is discrete).
\end{rm-lemma}
\begin{proof}
  We use a corollary that if there is a split-surjection from \(A\) to \(B\),
  and \(A\) is discrete, then \(B\) is also discrete.
\end{proof}
By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998}, we know that
this implies split enumerable types must all be sets.
\subsection{Closure}
In this section we will prove closure under various operations for split
enumerable sets.
We are working towards a topos proof, which requires us to prove closure under
a variety of operations: for now, we only have enough machinery to demonstrate
the semiring operations, and dependent sums.
In order to show closure under exponentials (function arrows), we will need an
equivalence with \(\mathbf{Fin}\), which will be provided in
section~\ref{manifest-bishop}.
\begin{rm-lemma} \label{split-enum-bool-top-bot}
  \(\mathbf{Bool}\), \(\top\), and \(\bot\) are all split enumerable.
\end{rm-lemma}
\begin{proof}
  These non-recursive types have similar, simple proofs.
  \begin{multicols}{3}
    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool}
    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{unit}
    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{empty}
  \end{multicols}
\end{proof}
\begin{rm-theorem} \label{split-enum-sigma}
  Split-enumerability is closed under \(\Sigma\).
  \begin{equation} \label{split-enum-sigma-clos-sig}
    \frac{
      \mathcal{E}!(A) \; \; \; \Pi (x : A) , \mathcal{E}!(U(x))
    }{
      \mathcal{E}!(\Sigma (x : A) , U(x))
    }
  \end{equation}
\end{rm-theorem}
\begin{proof}
  To obtain the support list, we concatenate the support lists of all the proofs
  of split-finiteness for \(U\) over the support list of \(E_A\).
  In Agda:
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sigma-enum}
  ``do-notation'' is available to us as we're working in the list monad.
\end{proof}
Closure under disjoint union and Cartesian product both follow from \(\Sigma\),
as these types can be defined in terms of \(\Sigma\).
\begin{align}
  A \times B &= \Sigma {(x : A)} , B \\
  A + B &= \Sigma {(x : \mathbf{Bool})} , \text{if} \; x \; \text{then} \; A \; \text{else} \; B
\end{align}
\section{Manifest Bishop Finiteness} \label{manifest-bishop}
Where split enumerability was the enumeration form of a split surjection from
\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
\emph{equivalence} with \(\mathbf{Fin}\).

@@

A nice consequence of prohibiting duplicates is that now the length of the
support list is the same as the cardinality of the set.
\subsection{Equivalence}
\begin{rm-lemma} \label{bishop-equiv}
  A proof of manifest Bishop finiteness is equivalent to an equivalence with a
  finite prefix of the natural numbers.
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{rm-lemma}
\begin{proof}
  There are many equivalent definitions of equivalence in HoTT.
  Here we take the version preferred in the Cubical Agda library: contractible
  maps \cite[definition 4.4.1]{hottbook}.
  Because of the parallels between contractible maps and split surjections,
  the proof proceeds much the same as \ref{split-enum-surj}.
  In other words, the definition of Bishop finiteness is itself a reassociation
  of a contractible map.
\end{proof}
\subsection{Relationship to Split Enumerability}
We now show that manifest Bishop finiteness has equal strength to split
enumerability.
\begin{rm-theorem} \label{split-enum-to-manifest-bishop}
  Any split enumerable set is manifest Bishop finite.
\end{rm-theorem}
\begin{proof}
  From proposition~\ref{split-enum-discrete} we can derive decidable equality on
  \(A\), and using this we can define a function (called \(\text{uniques}\)) which filters
  out duplicates from lists of \(A\)s.
  This gives us our support list.
  To generate the cover proof it suffices now to prove the following:
  \begin{equation}
    \Pi {(x : A)} , \Pi {(\mathit{xs} : \textbf{List}(A))} , x \in \mathit{xs} \rightarrow x \inunique \text{uniques}(\mathit{xs})
  \end{equation}
\end{proof}
\subsection{Closure}
Proving equal strength of split enumerability and manifest Bishop finiteness
allows us to carry all of the previous proofs of closure over to manifest Bishop
finite sets (and vice-versa).
Missing from our previous proofs was a proof of closure of functions.
We remedy that here.
\begin{rm-theorem}
  Manifest bishop finiteness is closed over dependent functions
  (\(\prod\)-types).
  \begin{equation}
    \frac{
      \mathcal{B}(A) \; \; \; \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
    }{
      \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
    }
  \end{equation}
\end{rm-theorem}
\begin{proof}
  This proof is essentially the composition of two transport operations, made
  available to us via univalence.

@@

@@

@@

@@

  Kuratowski finite sets are closed under \(\times\) and \(+\).

@@

  These follow from their lifted versions on manifestly enumerable types.

@@

In the previous sections we saw different flavours of finiteness which were
really just different flavours of relations to \(\mathbf{Fin}\).
In this section we will see that we can construct a similar classification of
relations to \(\mathbb{N}\), in the form of the countably infinite types.

@@

have a support \emph{stream}, which is its infinite.

@@

  As with lists, it is cleaner, theory-wise, to define streams as a container.

@@

    \mathbf{Stream}(A) = \llbracket \top \triangleright \text{const} (\mathbb{N}) \rrbracket(A) \simeq \mathbb{N} \rightarrow A

@@

@@

@@

  As mentioned, we will have to use a more sophisticated pairing function than
  the Cartesian product we used before.
  We instead will mirror the pattern in figure~\ref{cantor}.
  To greatly simplify the algorithm, we will produce an intermediate stream of
  lists which consists of the diagonals in the diagram.
  We then concatenate these streams into the final support stream.
\end{proof}
\begin{rm-lemma}
  Split countability is closed under non-dependent product and sum.
\end{rm-lemma}
\begin{proof}
  Follows from theorem~\ref{split-countability-sigma}.
\end{proof}
\subsubsection{Kleene Star}
While we lose some closures with the inclusion of infinite types, we gain some
others.
In particular, we have the Kleene star.
This means, in effect, that we have closure under lists.
\begin{rm-theorem}
  Split countability is closed under Kleene star.
  \begin{equation}
    \mathcal{E}!(A) \rightarrow \mathcal{E}!(\mathbf{List}(A))
  \end{equation}
\end{rm-theorem}
\begin{proof}
  As with the proof of closure under \(\Sigma\), our main task here is to figure
  out a way to arrange the indices such that 
\end{proof}
\subsection{Manifest Countability}
As we can quotient out the position information with finite types, so can we with
countable types.
\todo[inline]{fill in rest here}
\section{Practical Uses} \label{practical}
\subsection{Omniscience}
In this section we are interested in restricted forms of the limited principle
of omniscience \cite{myhillErrettBishopFoundations1972}.
\begin{rm-definition}[Limited Principle of Omniscience]
  For any type \(A\) and predicate \(P\) on \(A\), the limited principle of
  omniscience is as follows:
  \begin{equation}
    \left( \Pi {(x : A)} , \mathbf{Dec}(P(x)) \right) \rightarrow \mathbf{Dec} \left( \Sigma {(x : A)} , P(x) \right)
  \end{equation}
  In other words, for any decidable predicate the existential quantification of
  that predicate is also decidable.
\end{rm-definition}
The limited principle of omniscience is non-constructive, but individual types
can themselves satisfy omniscience.
In particular, \emph{finite} types are omniscient.

@@

Omniscience is stronger than exhaustibility, as we can derive the latter from
the former:
\begin{rm-lemma} \label{omniscient-is-exhaustible}
  Any omniscient type is exhaustible.
\end{rm-lemma}
\begin{proof}
  For decidable propositions, we know the following:
  \begin{equation} \label{lem-conversion}
    \Pi (x : A) , P(x) \leftrightarrow \neg \Sigma (x : A) , \neg P(x) 
  \end{equation}
  To derive exhaustibility from omniscience, then, we run the predicate in its
  negated form, and then subsequently negate the result.
  The resulting decision over \(\neg \Sigma(x : A) , \neg P(x)\) can be
  converted into \(\Pi (x : A) , P(x)\).
\end{proof}
We cannot derive, however, that any exhaustible type is omniscient, as we do not
have the inverse of equation~\ref{lem-conversion}:
\begin{equation} \label{choice-conversion}
  \Sigma (x : A) , P(x) \leftrightarrow \neg \Pi (x : A) , \neg P(x)
\end{equation}
Such an equation would allow us to pick a representative element from any type,
which is therefore non-constructive.
In a sense, equation~\ref{lem-conversion} requires a form of LEM on the
proposition (i.e. requires it to be decidable), whereas
equation~\ref{choice-conversion} requires a form of choice.
Those finiteness predicates which are ordered do in fact give us this form of
choice, so the conversion is valid.
As such, all of the ordered finiteness predicates imply omniscience.
Again, we will prove it only for the weakest.
\begin{rm-theorem}
  Manifest enumerable types are omniscient.
\end{rm-theorem}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}
Finally, we do have a form of omniscience for prop-valued predicates, as they do
not care about the chosen representative.
\begin{rm-theorem}
  Kuratowski finite types are omniscient about prop-valued predicates.
\end{rm-theorem}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}
\subsection{Synthesising Pattern-Matching Proofs}
In particular, they can automate large proofs by analysing every possible case.
In \cite{firsovDependentlyTypedProgramming2015}, the \(\AgdaDatatype{Pauli}\)
group is used as an example.
\ExecuteMetaData[agda/Data/Pauli.tex]{def}
As \(\AgdaDatatype{Pauli}\) has 4 constructors, \(n\)-ary functions on
\(\AgdaDatatype{Pauli}\) may require up to \(4^n\) cases, making even simple
proofs prohibitively verbose.

The alternative is to derive the things we need from \(\mathcal{E}!\) somehow.
As \(\AgdaDatatype{Pauli}\) is a simple finite type, the instance can be defined
in a similar way to those in lemma~\ref{split-enum-bool-top-bot}.
From here we can already derive decidable equality, a function which requires 16
cases if implemented manually.

For proof search, the procedure is a well-known one in Agda
\cite{devrieseBrightSideType2011}: we ask for the result of a decision procedure
as an \emph{instance argument}, which will demand computation during
typechecking.
As an example, we can check that the group operation on \(\AgdaDatatype{Pauli}\)
is idempotent like so:
\ExecuteMetaData[agda/Data/Pauli.tex]{noninst-prf}

As a quick aside, when the property is \emph{not} true (for instance \(\forall x
\rightarrow x \cdot x \equiv x\)), Agda will fail by not finding an instance of
\(\bot\).
The error message specifically is
\verb+No instance of type ⊥ was found in scope+, which isn't very helpful.
We can do a little better by displaying the particular element on which the
property failed, by defining a custom empty type parameterised by the
counterexample itself.
Combined with some other changes to the combinators, this will give the
following more helpful error message:
\verb+No instance of type Counterexample X was found in scope+.
\subsection{Multiple Arguments}
The automation machinery above only deals with single-argument predicates.
This is not a problem, as we know that we can work with multiple arguments by
currying and uncurrying, since all of the finiteness predicates are closed under
\(\times\).
To automate away the curry/uncurry noise we will use instance search, building
on \cite{allaisGenericLevelPolymorphic2019} to develop a small interface to
generic \(n\)-ary functions and properties.
Our generic representation can handle dependent \(\Sigma\) and \(\prod\)
types (rather than their non-dependent counterparts, \(\times\) and
\(\rightarrow\)).
This extension was necessary for our use case: it is mentioned in the paper
as the obvious next step. 
We also implement the curry-uncurry combinators as (verified) isomorphisms.

A full explanation of our implementation is beyond the scope of this work, so we
only present the finished interface, which is used like so:
\ExecuteMetaData[agda/Data/Pauli.tex]{prf-curried}

\todo[inline]{Synthesise functions}
\todo[inline]{Partial search over infinite types}
\bibliographystyle{splncs04}
\bibliography{bibliography.bib}
\end{document}
